---
title: "Week 8 - Input / Output"
output:
  html_notebook: default
---
```{r setup}
knitr::opts_knit$set(root.dir = "D:/Introduction-to-R/Data")
```
In previous weeks, the course briefly discussed the import and export of data. This week will take an even more in-depth look at the ability for R to import and export data files. This includes both local copies of data as well as access to online sources.

This week's assignment is about creating a farming simulation.

## Weekly Learning Objectives
Create an R script with the following components:

1. Build a simple R program to read in data from a file and save the results into a data frame.
2. Build a function to read data from a web-URL.

- Use the public archive data on Relative CPU performance: 
- URL: http://archive.ics.uci.edu/ml/machine-learning-databases/
- Dataset: machine.data
- Name File: machine.names

3.  Build a simple lookup function for CPU performance. Assume the data will be pulled from the same URL, and that data is updated on a frequent basis. Given a vendor name turn the following results:

- A message if the vendor is not found (based on previous experience in the data)
- Vendor published relative performance (PRP) and estimated relative performance (ERP)
- If the vendor is the best in category, return a flag that they are the best. For example, "Best in Published Relative Performance".
- If they are not the best in category, return the vendor is the best including the PRP and ERP numbers.


## 1. Read data into a data frame
This is simply a load into a data frame.  There are multiple ways of solving this question.

#### Full Points
Load data into a data frame.  You may also want to print off the results and use the class() function to check to make sure it's a data frame.  
```{r Read_DataFrame}
sample_dataframe <- read.csv("sample.txt", header = TRUE,sep="\t")    #  First row is a header.  
                                                                      #  Field separated by a tab (\t)
class(sample_dataframe)    # If it worked it will say data.frame
```
## 2. Read data from a web URL
This simply extends the function to pull from a web address.  Here I will use the 

#### Full Points
Load data into a data frame from a web link.  You may also want to print off the results and use the class() function to check to make sure it's a data frame.  

#### Red Wine Data
```{r Red_Wine}
RED_URL <-"http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv"
RED_data <- read.csv(RED_URL, header = TRUE, sep = ";")                                                             
head(RED_data)
```
#### White Wine Data
```{r White_Wine}
WHITE_URL <-"http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv"
WHITE_data <- read.csv(WHITE_URL, header = TRUE, sep = ";")  
head(WHITE_data)
```
## 3. Lookup Function
This comes down to a simple sort
- A message if it is not found
- Top record / bottom record
- Best in category

```{r Lookup_Function}

sorted_red <- RED_data[order(11),]
sorted_red
```

