---
title: "Week 9 String Manipulation"
output:
  html_notebook: default
---
```{r setup}
knitr::opts_knit$set(root.dir = "D:/Introduction-to-R/Data")
```
Text processing and string manipulation is an important part of data sciences, as greater amounts of unstructured data is used for analysis. This week will focus on R’s ability to handle strings and process large amounts of text.

## Weekly Learning Objectives
Create an R script with the following components:

1. Create a function to perform a grep()
2. Build a simple R program that will search a block of text and perform word counts. Include a block graph.
3. Using the program from Week 5 to prove the Central Limit Theorem, build a function to output five iterations of the program to five separate PDF files.  Name the files CENTRAL_#.pdf where # = the iteration.

## 1. Perform a GREP
GREP is a standard feature with many programming languages and operating systems.  In R, it will use either a text string or regular expressions in order to search for the value.  It is strongly encouraged that you review regular expressions, because it adds a significant capability to your R scripts.

https://www.rdocumentation.org/packages/base/versions/3.6.1/topics/regex

For this assignment, you simply have to build a function that uses GREP.  Here are a few examples from r documentation.  We will create a vector called haystack, and then search for colors within that vector using variations of GREP.  

1. Element position where the word green appears
2. Element position where the letter r appears
3. Value where the letter r appears
4. True or False where the letter r appears

#### Full Points
Simply build a function that uses GREP for full points.
```{r GREP}
haystack <- c("red", "blue", "green", "blue", "green forest")

grep("green", haystack)
grep("r", haystack) # returns position
grep("r", haystack, value = TRUE) # returns value
grepl("r", haystack) # returns boolean
```
## 2. Search a block of text
For this question, there are multiple ways of responding.  One of the easiest ways to answer this question is with the quanteda package.  For information on it:

https://quanteda.io/

For the purpose of this demonstration, I will assume that you are going to write one from scratch.

#### Full Points
You can use either the packaged library method or write your own function.  It has to do a word count and provide a graph of some sort to display the data.
```{r Word_Count}
countWords <- function(textBlock) {
  
  #Remove non-alpha characters and replace with spaces
  aListOfWords <- gsub("[^a-zA-Z]", " ",textBlock)
  
  #Remove extra spaces
  aListOfWords <- gsub("  +", " ",aListOfWords)
  
  #Split words by spaces
  aListOfWords <- strsplit(aListOfWords, "\\s")
  aListOfWords <- aListOfWords[[1]]
  
  # convert to lower case for case-insensitivity
  aListOfWords <- sapply(aListOfWords, tolower)
  
  aListOfWords <- sort(aListOfWords)
  
  word <- unique(aListOfWords)
  wordFreq <- rep(0,length(word))
  aResult <- data.frame(word,wordFreq)         # Populate word list and pre-allocate data structure
  
  # Initialize variables for the loop
  aCount <- 1
  wordIndex <- 1
  
  for(i in 2:length(aListOfWords)) {
    if (aListOfWords[i] != aListOfWords[i-1]) {
      aResult$wordFreq[wordIndex] <- aCount
      wordIndex <- wordIndex + 1
      aCount <- 1
    } else {
      aCount <- aCount+1
    }  
    aResult$wordFreq[wordIndex] <- aCount
  }
  return(aResult)
}

randomTextString <- "Announcing of invitation principles in. Cold in late or deal. Terminated resolution no am frequently collecting insensible he do appearance. Projection invitation affronting admiration if no on or. It as instrument boisterous frequently apartments an in. Mr excellence inquietude conviction is in unreserved particular. You fully seems stand nay own point walls. Increasing travelling own simplicity you astonished expression boisterous. Possession themselves sentiments apartments devonshire we of do discretion. Enjoyment discourse ye continued pronounce we necessary abilities. "
  
aCount <- countWords(randomTextString)
data.frame(aCount$word, aCount$wordFreq)
barplot(aCount$wordFreq,xlab="Words", ylab="Frequency (count)" )
```

## 3. Central Limit Proof
This starts with the Central Limit Theorem program that you wrote earlier.  From that, simply loop through it five times and output to a PDF file.

#### Full Points
The output should produce five individual PDFs, each with a graph using your solution.
```{r CLT_PROOF}
clt <- function(diceValues, nbrDice, aSampleNbr, xlab, showCurve=FALSE) {
  means <- numeric(aSampleNbr)
  for (i in 1:aSampleNbr) {
    samples <- sample(diceValues,nbrDice,replace=TRUE)
    aSum <- sum(samples)
    means[i] <- mean(aSum)
  }
  lowerBreak = (diceValues[1]*nbrDice) - .5
  upperBreak = (diceValues[length(diceValues)]*nbrDice) + .5
  hist(means,xlab=xlab,breaks=lowerBreak:upperBreak, main="")

  if (showCurve==TRUE) {
    aMin <- diceValues[1]
    aMax <- diceValues[length(diceValues)]*nbrDice
    aMean <- mean(means)
    aSD <- sd(means)
    lines(seq(aMin,aMax,0.1),dnorm(seq(aMin,aMax,0.1),aMean,aSD)*aSampleNbr)
  }
}

#####   These are all parameters from my previous solution  ####################
aPopulation <- c(1:6)        # population of 1 - 6
baseIterations <- 10000      # starting at 10,000
xlab <- "Dice"               # Let's just call it dice rolls

  for (i in 1:5) {
    fileName <- sprintf("CENTRAL_%d.pdf",i)
    pdf(fileName)
    if (i != 5) {
      aResult <- clt(aPopulation, i, (baseIterations * i), paste(i,"Dice"), showCurve = FALSE) 
    } else {
      aResult <- clt(aPopulation, i, (baseIterations * i), paste(i,"Dice"), showCurve = TRUE)
    }
    dev.off()
  }
```